Finite element analysis of multi-dimensional and simplified models for beams

### 2024-04-15 final_seminar/main.tex
order change and added datat

\documentclass[8pt]{beamer}
\usetheme{metropolis} % Use the Metropolis theme
\usepackage{pgfplots}
\usepackage{caption,subcaption}
\usepackage[style=authoryear,backend=biber]{biblatex}
\addbibresource{../references.bib}

@@

    \subsection{Existence and uniqueness of solutions}

        In [\cite{VV02}] the authors present results for the existence and uniqueness of solutions to the general vibration problem. The following assumptions are made for the existence results.

        \begin{frame}{Existence and uniqueness of solutions}
            The following assumptions are made in \cite{VV02} for the existence results.
            \begin{itemize}
                \item[] \textbf{A1} - $V$ is dense in $W$ and $W$ is dense in $X$.

            \only<2>{
            Dividing both sides by $T(t)$ gives
            \begin{eqnarray*}
                b(x,v) = - \frac{T''(t)}{T(t)}c(x,v) \implies  \frac{T''(t)}{T(t)} \ \ \ \textrm{must be constant.}
            \end{eqnarray*}}
        \end{frame}

        \begin{frame}{Modal Analysis}
            Set $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$.

            At this point we do not yet know if such a $\lambda$ exists.

            \only<2->{
            Eigenvalue Problem:

            Find a real number $\lambda$ and a $x \in V$ with $x \neq 0$ such that
            \begin{eqnarray*}
                b(x,y) = \lambda c(x,y) \ \ \ \ \textrm{ for each } \ y \in V.
            \end{eqnarray*}}

@@

         \begin{frame}{Modal Analysis}
            \only<1->{
            Eigenvalue Problem:

            Find a real number $\lambda$ and a $x \in V$ with $x \neq 0$ such that
            \begin{eqnarray*}
                b(x,y) = \lambda c(x,y) \ \ \ \ \textrm{ for each } \ y \in V.
            \end{eqnarray*}}

            Existence of a solution is presented in the article [\cite{CVV18}]\footfullcite{CVV18}

@@

        \begin{frame}{Modal Analysis}
            \begin{itemize}
                \item[] \textbf{A1} - $V$ is dense in $W$ and $W$ is dense in $X$.

                \item[] \textbf{A5} - The embedding of $V$ into $W$ is compact.
            \end{itemize}
         \end{frame}

         \begin{frame}{Modal Analysis}
            Formal series solution for the general vibration problem:
            \begin{eqnarray}
                u(t) = \sum_{n=1}^{\infty} T_n(t)x_n. \label{eq:1D_Model:ModalAnalysisSeriesSolution}
            \end{eqnarray}
            with 
            \begin{flalign}
                T_n(t) &=  A_n \cos(\sqrt{\lambda_n} t) + B_n \sin(\sqrt{\lambda_n} t) \quad \textrm{ if } \lambda_n > 0, \label{lambda_1}
            \end{flalign}
         \end{frame}

@@

\section{Literature study}
    \subsection{Comparison of linear beam theories}
        \begin{frame}{Literature study - Comparison of linear beam theories}
            Consider the article [\cite{LVV09}]\footfullcite{LVV09}
        \end{frame}

    \subsection{Empirical comparison of beams}
        \begin{frame}{Literature study - Empirical comparison of beams}
            Consider the article [\cite{SP06}]\footfullcite{SP06}
        \end{frame}

@@

>>>>

\documentclass[8pt]{beamer}
\usetheme{metropolis} % Use the Metropolis theme
\usepackage{pgfplots}
\usepackage{caption,subcaption}
\usepackage[style=authoryear,backend=biber]{biblatex}
\usepackage[table]{xcolor}
\addbibresource{../references.bib}

@@

\section{Literature study}
    \subsection{Comparison of linear beam theories}
        \begin{frame}{Literature study - Comparison of linear beam theories}
            Consider the article \textit{Comparison of linear beam theories} [\cite{LVV09}]

            \footfullcite{LVV09}
        \end{frame}

        \begin{frame}{Detailed Comparison: Timoshenko Beam Model vs. Two-Dimensional Model}

            \textbf{Objective of Comparison:}
            \begin{itemize}
              \item To validate the efficacy of the Timoshenko beam theory against a more comprehensive two-dimensional model, which accounts for plane stress conditions and detailed stress distribution across the beam's cross-section.
            \end{itemize}

            \textbf{Methodology:}
            \begin{itemize}
              \item Utilized finite element methods to calculate natural frequencies and modes, providing a direct comparison of results from both models.
              \item Both models analyzed under the same boundary conditions and loading scenarios to ensure consistency in comparison.
            \end{itemize}

            \textbf{Key Findings:}
            \begin{itemize}
              \item \textbf{Close Agreement}: Timoshenko model results closely match those from the two-dimensional model, with most eigenvalues showing less than 3\% deviation.
              \item \textbf{Model Reliability}: Demonstrates high reliability of the Timoshenko model for predicting beam behavior under various loading conditions, particularly for lower vibration modes.
            \end{itemize}

@@

        \begin{frame}
            \textbf{Model Advantages:}
            \begin{itemize}
              \item \textbf{Timoshenko Model}: Offers significant improvements in accuracy over simpler models like Euler-Bernoulli, especially in scenarios where beam lengths are relatively short and shear deformations cannot be ignored.
              \item \textbf{Two-Dimensional Model}: Provides a more rigorous analysis by accounting for two-dimensional stress effects, which are important for comprehensive structural analyses.
            \end{itemize}

            \textbf{Practical Implications:}
            \begin{itemize}
              \item \textbf{Engineering Applications}: Timoshenko model is particularly advantageous for practical engineering applications, providing a good balance between computational efficiency and accuracy.
              \item \textbf{Advanced Analyses}: For critical engineering tasks where detailed stress distribution is crucial, the two-dimensional model serves as a more suitable analytical tool.
            \end{itemize}

            \textbf{Conclusions and Recommendations:}
            \begin{itemize}
              \item \textbf{Preferred Use of Timoshenko Model}: Recommended for most practical engineering applications involving beams, where it strikes an optimal balance between accuracy and simplicity.
              \item \textbf{When to Use Two-Dimensional Model}: Advisable for in-depth structural analyses where precise stress distribution is required, such as in complex assemblies or where materials with varying properties are involved.
              \item This comparison serves as an intermediate step towards the more comprehensive analysis involving the comparison of two-dimensional and three-dimensional models.
            \end{itemize}

@@

    \subsection{Empirical comparison of beams}
        \begin{frame}{Literature study - Empirical comparison of beams}
            Consider the article \textit{On the valid frequency range of Timoshenko beam theory} [\cite{SP06}]\footfullcite{SP06}
        \end{frame}

        \begin{frame}{Setup and Testing Method}
            \textbf{Objective:}
            \begin{itemize}
                \item To validate the frequency predictions of Timoshenko beam theory (TBT) for different end conditions and frequencies.
            \end{itemize}

            \textbf{Methodology:}
            \begin{itemize}
                \item Experimental setup involved a short free-free beam of near-square cross-section.
                \item Beam dimensions: Length = 40.047 mm, Breadth = 10.097 mm, Depth = 10.047 mm.
                \item Beam material: Aluminum alloy, with properties determined via resonant ultrasound spectroscopy (RUS).
                \item Comparisons made with experimental observations, commercial finite element (ANSYS) predictions, and RUS technique predictions.
                \item Finite element analysis used 20-node brick elements, and TBT predictions were calculated using specific frequency equations.
            \end{itemize}

            \textbf{Key Setup Features:}
            \begin{itemize}
                \item The experimental beam was suspended and excited by two carbon-fibre loops attached to piezoelectric transducers.
                \item Resonant frequencies detected by sweeping the frequency range with a network analyzer.
            \end{itemize}
        \end{frame}

        \begin{frame}{Numerical Results and Discussion}
            \textbf{Results Summary:}
            \begin{itemize}
                \item Excellent agreement between measured, ANSYS, and RUS frequencies below the cut-off frequency.
                \item Above the cut-off frequency, TBT predictions deviate increasingly from measured and simulated frequencies.
            \end{itemize}

            \textbf{Discussion:}
            \begin{itemize}
                \item For free-free end conditions, discrepancies arise above the cut-off frequency, suggesting limitations in the TBT's applicability.
                \item TBT predictions were increasingly inaccurate for higher modes (n >= 8), exceeding the typical engineering tolerance.
            \end{itemize}

            \textbf{Conclusion:}
            \begin{itemize}
                \item The study suggests a cut-off frequency as a reasonable upper bound for TBT predictions.
                \item This research serves as an intermediate step, setting the stage for further comparison between two-dimensional and three-dimensional models.
            \end{itemize}
        \end{frame}

@@

    \subsection{Existence and uniqueness of solutions}
        \begin{frame}{Existence and uniqueness of solutions}
            In [\cite{VV02}] the authors present results for the existence and uniqueness of solutions to the general vibration problem. The following assumptions are made for the existence results.

            \begin{itemize}
                \item[] \textbf{A1} - $V$ is dense in $W$ and $W$ is dense in $X$.

            \only<2>{
            Dividing both sides by $T(t)$ gives
            \begin{eqnarray*}
                b(x,v) = - \frac{T''(t)}{T(t)}c(x,v) \implies  \frac{T''(t)}{T(t)} \ \ \ \textrm{must be constant.}
            \end{eqnarray*}}

            \only<3->{
                Set $\displaystyle \frac{T''(t)}{T(t)} = -\lambda$.

            At this point we do not yet know if such a $\lambda$ exists.
            }
        \end{frame}

        \begin{frame}{Modal Analysis}
            Eigenvalue Problem:

            Find a real number $\lambda$ and a $x \in V$ with $x \neq 0$ such that
            \begin{eqnarray*}
                b(x,y) = \lambda c(x,y) \ \ \ \ \textrm{ for each } \ y \in V.
            \end{eqnarray*}

            Ordinary differential equation for $T(t)$:
            \begin{eqnarray}
                T''(t)  + \lambda T(t) = 0. \label{eq:1D_Model:ModalAnalysisODE}
            \end{eqnarray}
         \end{frame}

        \begin{frame}{Modal Analysis}
            Assumptions:
            \begin{itemize}
                \item[] \textbf{A1} - $V$ is dense in $W$ and $W$ is dense in $X$.

                \item[] \alert{\textbf{A5} - The embedding of $V$ into $W$ is compact.} [\cite{CVV18}]
            \end{itemize}
            \footfullcite{CVV18}
        \end{frame}

        \begin{frame}{Modal Analysis}

            ***************Verduideliking********************

            Solutions for the Eigenvalue Problem:
            \begin{eqnarray*}
                u = \sum_{i=1}^{\infty} a_i x_i = \sum_{i=1}^{\infty} c(u, x_i)x_i.
            \end{eqnarray*}

            Solutions for the Ordinary Differential Equation:
            \begin{flalign}
                T_n(t) &=  A_n \cos(\sqrt{\lambda_n} t) + B_n \sin(\sqrt{\lambda_n} t) \quad \textrm{ if } \lambda_n > 0, \label{lambda_1}
            \end{flalign}

            Formal series solution for the general vibration problem:
            \begin{eqnarray}
                u(t) = \sum_{n=1}^{\infty} T_n(t)x_n.
            \end{eqnarray}
        \end{frame}

@@

        \begin{frame}
            \begin{table}[htbp]
                \scalebox{.8}{
                \makebox[\textwidth]{
                    \begin{tabular}{|cc|cc|cc|cc||cc|}
                        \hline
                        \multicolumn{10}{|c|}{Eigenvalues for $b \leq h$} \\
                        \hline
                        \hline
                        i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
                        \hline
                        {2} & 0.12307 & {2} & 0.12234 & {2} & 0.12198 & {3} & 0.12178 & 1     & 0.12151 \\
                        {3} & 3.5773 & {5} & 3.5630 & {6} & 3.5558 & {8} & 3.5519 & 2     & 3.5460 \\
                        \rowcolor{lightgray}{5} & 7.7799 & {6} & 7.7596 & {8} & 7.7471 & {11} & 7.7401 & 3     & 7.7311 \\
                        {8} & 20.334 & {9} & 20.283 & {11} & 20.26 & {14} & 20.247 & 4     & 20.225 \\
                        {10} & 56.247 & {12} & 56.173 & {15} & 56.156 & {22} & 56.142 & 5     & 56.109 \\
                        \rowcolor{lightgray}{11} & 69.197 & {14} & 69.319 & {17} & 69.281 & {24} & 69.238 & 6     & 69.164 \\
                        {14} & 114.03 & {16} & 114.01 & {20} & 114.05 & {29} & 114.06 & 7     & 114.03 \\
                        \rowcolor{lightgray}{17} & 187.01 & {19} & 189.14 & {25} & 189.37 & {36} & 189.34 & 8     & 189.17 \\
                        {18} & 192.21 & {20} & 192.41 & {26} & 192.58 & {37} & 192.63 & 9     & 192.61 \\
                        {21} & 284.76 & {23} & 285.43 & {31} & 285.74 & {42} & 285.84 & 10    & 285.85 \\
                        {23} & 327.57 & {26} & 328.24 & {35} & 328.37 & {46} & 328.40 & 11    & 328.40 \\
                        \rowcolor{lightgray}{25} & 347.77 & {28} & 356.44 & {36} & 357.30 & {50} & 357.33 & 12    & 357.08 \\
                        {27} & 393.69 & {30} & 396.84 & {38} & 397.28 & {53} & 397.37 & 13    & 397.33 \\
                        {30} & 434.46 & {34} & 441.05 & {41} & 441.81 & {57} & 441.99 & 14    & 442.00 \\
                        \rowcolor{lightgray}{31} & 523.65 & {36} & 534.04 & {43} & 534.17 & {63} & 534.03 & 15    & 533.71 \\
                        {34} & 550.51 & {37} & 537.82 & {44} & 538.86 & {64} & 539.06 & 16    & 538.97 \\
                        \rowcolor{lightgray}{37} & 590.86 & {41} & 587.43 & {48} & 594.17 & {65} & 595.58 & 17    & 596.06 \\
                        {39} & 590.86 & {42} & 600.52 & {49} & 602.25 & {67} & 602.69 & 18    & 602.77 \\
                        \rowcolor{lightgray}{42} & 646.21 & {44} & 657.22 & {50} & 658.04 & {71} & 658.06 & 19    & 657.87 \\
                        {44} & 711.07 & {46} & 714.62 & {53} & 717.10 & {73} & 717.51 & 20    & 717.37 \\
                        \hline
                        \hline
                        \multicolumn{2}{|c|}{Max RE: \  2.6069\%} &\multicolumn{2}{c|}{Max RE: \ 1.4469\%}  & \multicolumn{2}{c|}{Max RE: \  0.38192\%}  & \multicolumn{2}{c||}{Max RE: \ 0.22301\%}& \multicolumn{2}{c|}{-} \\
                        \hline
                    \end{tabular}%
                    \label{tab:2v3_1}%
                }}
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \centering
                \begin{tabular}{|c|cccc|}
                    \hline
                    \multicolumn{5}{|c|}{Maximum relative error for beam type and non-beam type eigenvalues for $h = 1/5$.} \\
                    \hline
                    \hline
                    & {b = h} & {b = 1/2h} & {b = 1/4h} & {b = 1/8h} \\
                    \hline
                    Beam Type & 2.1420 \% & 0.6804 \% & 0.38192 \% & 0.22301 \% \\
                    Non-Beam Type & 2.6069 \% & 1.4469 \% & 0.31546 \% & 0.11680 \% \\
                    \hline
                \end{tabular}%
                \label{tab:2Dv3D_1_breakup}%
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[ht]
                \scalebox{.8}{
                \makebox[\textwidth]{
                \begin{tabular}{|cc|cc|cc|cc||cc|}
                    \hline
                    \multicolumn{10}{|c|}{Comparsion of Eigenvalues with $h = 1/20$, with decreasing $b$ and $b < h$.} \\
                    \hline
                    \hline
                    i     & {b = h} & i     & {b = 1/2 h} & i     & {b = 1/4 h} & i     & {b = 1/8 h} & j     & 2D \\
                    \hline
                    {2} & 0.008043 & {2} & 0.008029 & {2} & 0.008023 & {3} & 0.00802 & {1} & {0.008013} \\
                    {3} & 0.3087 & {4} & 0.30816 & {5} & 0.30794 & {7} & 0.30785 & {2} & {0.30757} \\
                    {5} & 2.3357 & {8} & 2.3316 & {9} & 2.3300  & {12} & 2.3293 & {3} & {2.3273} \\
                    \rowcolor{lightgray}{8} & 7.7217 & {10} & 7.7156 & {13} & 7.7124 & {16} & 7.7111 & {4} & {7.7077} \\
                    {10} & 8.5387 & {11} & 8.5238 & {14} & 8.5182 & {18} & 8.516 & {5} & {8.5086} \\
                    {11} & 21.986 & {14} & 21.948 & {18} & 21.934 & {24} & 21.929 & {6} & {21.911} \\
                    {14} & 45.863 & {18} & 45.781 & {21} & 45.756 & {30} & 45.746 & {7} & {45.712} \\
                    \rowcolor{lightgray}{17} & 69.444 & {19} & 69.408 & {25} & 69.385 & {33} & 69.374 & {8} & {69.344} \\
                    {18} & 83.149 & {22} & 82.999 & {27} & 82.960 & {36} & 82.944 & {9} & {82.887} \\
                    {21} & 136.44 & {25} & 136.19 & {31} & 136.14 & {42} & 136.12 & {10} & {136.03} \\
                    \rowcolor{lightgray}{23} & 192.62 & {27} & 192.62 & {35} & 192.58 & {47} & 192.56 & {11} & {192.48} \\
                    {25} & 207.87 & {29} & 207.5 & {36} & 207.43 & {48} & 207.41 & {12} & {207.29} \\
                    {27} & 299.14 & {33} & 298.63 & {41} & 298.56 & {55} & 298.53 & {13} & {298.38} \\
                    \rowcolor{lightgray}{30} & 376.68 & {35} & 377.01 & {44} & 377.01 & {58} & 376.98 & {14} & {376.83} \\
                    {31} & 411.58 & {37} & 410.89 & {46} & 410.83 & {61} & 410.81 & {15} & {410.63} \\
                    {34} & 546.15 & {40} & 545.27 & {50} & 545.24 & {66} & 545.23 & {16} & {545.03} \\
                    \rowcolor{lightgray}{37} & 620.77 & {42} & 622.02 & {53} & 622.19 & {69} & 622.18 & {17} & {621.95} \\
                    {39} & 703.55 & {44} & 702.49 & {54} & 702.29 & {71} & 702.53 & {18} & {702.31} \\
                    {42} & 884.27 & {47} & 883.02 & {59} & 883.14 & {82} & 883.19 & {19} & {882.96} \\
                    \rowcolor{lightgray}{44} & 923.68 & {49} & 926.88 & {60} & 927.43 & {86} & 927.50 & {20} & {927.18} \\
                    \hline
                    \hline
                    \multicolumn{2}{|c|}{Max RE: \  0.37701\%} &\multicolumn{2}{c|}{Max RE: \ 0.19893\%}  & \multicolumn{2}{c|}{Max RE: \  0.12393\%}  & \multicolumn{2}{c||}{Max RE: \ 0.092843\%}& \multicolumn{2}{c|}{-} \\
                    \hline
                \end{tabular}%
                \label{tab:2v3_2}%
            }}
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \centering
                \begin{tabular}{|c|cccc|}
                    \hline
                    \multicolumn{5}{|c|}{Maximum relative error for beam type and non-beam type eigenvalues for $h = 1/20$} \\
                    \hline
                    \hline
                    & {b = h} & {b = 1/2h} & {b = 1/4h} & {b = 1/8h} \\
                    \hline
                    Beam Type & 0.37701 \% & 0.19893 \% & 0.12393 \% & 0.092843 \% \\
                    Non-Beam Type & 0.37682 \% & 0.10218 \% & 0.061213 \% & 0.043601 \% \\
                    \hline
                \end{tabular}%
                \label{tab:2v3_2_split}%
            \end{table}%
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \scalebox{.8}{
                \makebox[\textwidth]{
                    \begin{tabular}{|cc|cc|cc||cc|}
                        \hline
                        \multicolumn{8}{|c|}{Comparsion of Eigenvalues with $h = 1/20$, with increasing $b$ and $b > h$} \\
                        \hline
                        \hline
                        i     & {b = 2h} & i     & {b = 4h} & i     & {b = 8h} & j     & 2D \\
                        \hline
                        {2} & 0.008076 & {1} & 0.008162 & {1} & 0.008324 & 1     & 0.008013 \\
                        {3} & 0.30995 & {3} & 0.31298 & {3} & 0.31738 & 2     & 0.30757 \\
                        {5} & 2.3462 & {5} & 2.3737 & {6} & 2.4116 & 3     & 2.3273 \\
                        \rowcolor{lightgray}{8} & 7.7312 & {8} & 7.7471 & {9} & 7.7711 & 4     & 7.7077 \\
                        {10} & 8.5841 & {9} & 8.7082 & {10} & 8.7929 & 5     & 8.5086 \\
                        {11} & 22.124 & {12} & 22.491 & {14} & 23.222 & 6     & 21.911 \\
                        {14} & 46.195 & {14} & 47.003 & {18} & 47.921 & 7     & 45.712 \\
                        \rowcolor{lightgray}{17} & 69.454 & {17} & 69.281 & {22} & 72.307 & 8     & 69.344 \\
                        {18} & 83.822 & {18} & 85.218 & {24} & 80.607 & 9     & 82.887 \\
                        {21} & 137.43 & {21} & 138.58 & {32} & 140.97 & 10    & 136.03 \\
                        \hline
                        \hline
                        \multicolumn{2}{|c|}{Max RE: \  1.1289\%} &\multicolumn{2}{c|}{Max RE: \ 2.8261\%}  & \multicolumn{2}{c|}{Max RE: \  5.9821\%} & \multicolumn{2}{c|}{-} \\
                        \hline
                    \end{tabular}%
                    \label{tab:b>h_20}%
                }}
            \end{table}
        \end{frame}

        \begin{frame}
            \begin{table}[htbp]
                \centering
                \begin{tabular}{|c|ccc|}
                    \hline
                    \multicolumn{4}{|c|}{Maximum relative error for beam type and non-beam type eigenvalues for $h = 1/20$} \\
                    \hline
                    \hline
                    & {b = 2h} & {b = 4h} & {b = 8h} \\
                    \hline
                    Beam Type & 1.1289 \% & 2.8261 \% & 5.9821 \% \\
                    Non-Beam Type & 0.30521 \% & 0.51096 \% & 4.2734 \% \\
                    \hline
                \end{tabular}%
                \label{tab:b>h-split_20}%
            \end{table}%
        \end{frame}
